Standardized crypto-loans on\\ the Cardano blockchain

### 2021-02-04 papers/actus2021/article.tex
fixes

@@

\noindent 
\begin{equation}
\ident{interestRate}_{t+1}=\ident{capfloor}(\ident{interestRate}_{t-1}+\Delta_{r}),
\end{equation}

@@

A counterparty might decide to reinvest profit received as interest
from the %same 
% SJT I think "same" is superfluous here: is that OK?
loan. In the simplest case, this renders as compound
interest. This can be modelled through interest accrual and capitalisation;
% SJT what does "capitalisation" mean?
for instance, contracts from the ACTUS specification accrue interest
between interest payments and can transfer interest to a notional
during interest capitalisation event (IPCL).

@@

Marlowe~\cite{marlowe} is a high-level DSL for writing
financial contracts on blockchain. Marlowe is defined
by an executable semantics in Haskell, and has been implemented on
the UTxO-based Cardano blockchain. 
% SJT added this
Marlowe contracts are finite, and moreover have defined finite lifetime. This means that it is possible to perform \emph{static analysis} on them. For example, it is possible symbolically to check \emph{all} execution paths without running a contract, and in particular checking that every \texttt{Pay} construct can be executed in full.

@@

% SJT - the "i" and "t" subscripts here look a bit odd: is that what's intended?
\noindent 
\begin{equation}
\ident{riskfactor}_{i}{}_{t}=O_{\footnotesize\ident{rf}}(i,t)
\end{equation}
\noindent
In the case of the Cardano blockchain, these values are usually
provided through an \emph{oracle} mechanism\cite{oracles}. An oracle
could be a trusted party providing necessary data or network of parties
under consensus {[}find a reference{]}. 
% SJT REF needed
From a Marlowe DSL perspective,
the exact mechanism that provides external data is less important,
as Marlowe abstracts over IO by requiring a particular type of input -- a 
\texttt{Choice} -- that is protected with a cryptographic signature by the source of the choice.
% SJT changed from "owner" to "source".
As a result, the event of receiving data from an oracle is treated the
same as receiving numeric input in other languages

@@

% SJT - do you mean \mapsto or \rightarrow here?
\noindent 
\begin{equation}
\ident{Applicability}:\ident{ContractTerms}\mapsto \ident{Bool}
\end{equation}

ACTUS defines a family of applicability functions polymorphic on contract
type (TODO link to ACTUS applicability):
% SJT something needed here.

\smallskip
% SJT changed ContractTerm to ContractTerms
\noindent 
\begin{equation}
\ident{Applicability} : \ident{ContractType} \times \ident{ContractTerms} \mapsto \ident{ApplicabilityType}
\end{equation}
\noindent
where applicability could be: \emph{none}, \emph{always}, \emph{nullable}, or \emph{multiple}.

@@

\noindent
where the component \emph{receiveData} asks an oracle for Marlowe Choice if needed, and
\mbox{\emph{calculatePayoff}} calculates the payoff formula. For fixed-rate contracts
this is optimised into a pre-calculated constant function. The \emph{processPayoff} function
awaits the \texttt{Deposit} of a payoff amount from a party: if the deposit is made them 
it directs the funds to a counterparty, otherwise it transfers
the collateral to the counterparty and closes the contract. The principal components of the system are shown in Figure \ref{fig1}.
% SJT added this figure ref: OK?
% Do we need to add a little more to explain Fig1?

The chain
is generated from the fixed schedule of events, known in advance of execution, 
using the \ident{SCHED()} function from ACTUS, as shown in Figure \ref{fig2}.
% SJT added this figure ref: OK?

@@

% In a long-term perspective
There is a possible workaround: contracts with no maturity date
could be represented as actors with a finite number of state transitions.
% SJT is this ^^^ rephrasing OK?
We prototyped this approach, % in a separate branch, 
however it does
seem more prone to errors comparing to rendering predefined schedules.
More importantly, it greatly affects static analysis because the number
of reduction steps in the contract grows from 
\noindent 
\begin{equation*}
N_{\footnotesize \ident{scheduled}}=\ident{count(event)}
\end{equation*}
to 
\noindent 
\begin{equation}
N_{\footnotesize \ident{stateTransitions}}=\frac{\ident{max(date(event))}-\ident{min(date(event))}}{\ident{precision}} 
\end{equation}

@@

For numeric types Marlowe supports Integers, while ACTUS is expressed in terms
of real numbers. In order to model a real number in such a setup
we rely on fixed-point precision. The algebra looks like this: 
\begin{verbatim}
  (+)   = AddValue
  (-)   = SubValue
  a * b = Scale (1 % marloweFixedPoint) $ MulValue a b
\end{verbatim}
% SJT this needs more explanation here ^^^
\noindent
We plan to move Marlowe to fixed-precision numbers for the on-blockchain implementation available later in 2021.

@@

The Marlowe DSL does not support any notion of records,
the variables can only be of type ``Integer''. In order to
map contract state -- \emph{ContractStatePoly} -- we pack a set of Marlowe
variables of type \texttt{Value} and \texttt{Observation}, representing the previous ($t-1$) state, passing them into
\emph{ContractStatePoly}, to apply the polymorphic state transition, and
finally unpack \emph{ContractStatePoly} into a set of Marlowe variables representing the 
 state at $t+1$:
% SJT again I am confused about t-1 to t+1 transition. Why not just one step?  
\noindent 
\begin{equation}
st_{t+1}=unpack(stateTransition(pack(st_{t-1}))),
\end{equation}
\noindent
where \emph{pack} is a chain of \texttt{UseValue} constructs and \ident{unpack} is a chain of
\texttt{Let}s.

>>>>

% dk14 
% A1 it is from the ACTUS spec - I assume "r" means "rate"
% A2 I meant "before vs after event", but I'll remove it

@@

\noindent 
\begin{equation}
\ident{interestRate}_{t}=\ident{capfloor}(\ident{interestRate}_{t-1}+\Delta_{r}),
\end{equation}

@@

A counterparty might decide to reinvest profit received as interest
from the %same 
% SJT I think "same" is superfluous here: is that OK?
% dk14 sure
loan. In the simplest case, this renders as compound
interest. This can be modelled through interest accrual and capitalisation;
% SJT what does "capitalisation" mean?
% dk14 "conversion of income or assets into capital" it's quite googlable financial terms so idk if I should note it explicitly
for instance, contracts from the ACTUS specification accrue interest
between interest payments and can transfer interest to a notional
during interest capitalisation event (IPCL).

@@

Marlowe~\cite{marlowe} is a high-level DSL for writing
financial contracts on blockchain. Marlowe is defined
by an executable semantics in Haskell, and has been implemented on
the UTxO-based Cardano blockchain. 
% SJT added this
%dk14 OK
Marlowe contracts are finite, and moreover have defined finite lifetime. This means that it is possible to perform \emph{static analysis} on them. For example, it is possible symbolically to check \emph{all} execution paths without running a contract, and in particular checking that every \texttt{Pay} construct can be executed in full.

@@

% SJT - the "i" and "t" subscripts here look a bit odd: is that what's intended?
% dk14 - fixed
\noindent 
\begin{equation}
\ident{riskfactor}_{it}=O_{\footnotesize\ident{rf}}(i,t)
\end{equation}
\noindent
In the case of the Cardano blockchain, these values are usually
provided through an \emph{oracle} mechanism\cite{oracles}. An oracle
could be a trusted party providing necessary data or network of parties
under consensus {[}find a reference{]}. 
% SJT REF needed
From a Marlowe DSL perspective,
the exact mechanism that provides external data is less important,
as Marlowe abstracts over IO by requiring a particular type of input -- a 
\texttt{Choice} -- that is protected with a cryptographic signature by the source of the choice.
% SJT changed from "owner" to "source".
As a result, the event of receiving data from an oracle is treated the
same as receiving numeric input in other languages

@@

% SJT - do you mean \mapsto or \rightarrow here?
% dk14 - it is a mapping, but could be morphism too
\noindent 
\begin{equation}
\ident{Applicability}:\ident{ContractTerms}\rightarrow \ident{Bool}
\end{equation}

ACTUS standard defines a family of applicability functions polymorphic on contract
type:
% SJT something needed here.
% dk14 - we already gave a link to actusfrf, so I think I can skip references

\smallskip
% SJT changed ContractTerm to ContractTerms
\noindent 
\begin{equation}
\ident{Applicability} : \ident{ContractType} \times \ident{ContractTerms} \rightarrow \ident{ApplicabilityType}
\end{equation}
\noindent
where applicability could be: \emph{none}, \emph{always}, \emph{nullable}, or \emph{multiple}.

@@

\noindent
where the component \emph{receiveData} asks an oracle for Marlowe Choice if needed, and
\mbox{\emph{calculatePayoff}} calculates the payoff formula. For fixed-rate contracts
this is optimised into a pre-calculated constant function. The \emph{processPayoff} function
awaits the \texttt{Deposit} of a payoff amount from a party: if the deposit is made them 
it directs the funds to a counterparty, otherwise it transfers
the collateral to the counterparty and closes the contract. The principal components of the system are shown in Figure \ref{fig1}. There are 3 categories of components representing ACTUS functions: specification with formulas, formula wiring for fixed rates (produces precomputed payoffs) and formula wiring for variable rates (produces Marlowe-code that computes payoffs). Different types of wiring correspond to different implementations of \emph{ActusOps} implemetations (either \emph{Double} or \emph{Value}). 
% SJT added this figure ref: OK?
% Do we need to add a little more to explain Fig1?
% dk14 OK, added few sentences about Fig1

The chain
is generated from the fixed schedule of events, known in advance of execution, 
using the \ident{SCHED()} function from ACTUS, as shown in Figure \ref{fig2}.
% SJT added this figure ref: OK?
% dk14 OK

@@

% In a long-term perspective
There is a possible workaround: contracts with no maturity date
could be represented as actors with a finite number of state transitions.
% SJT is this ^^^ rephrasing OK?
% dk14 - sure
We prototyped this approach, % in a separate branch, 
however it does
seem more prone to errors comparing to rendering predefined schedules.
More importantly, it greatly affects static analysis because the number
of reduction steps in the contract grows from 
\noindent 
\begin{equation*}
N_{\footnotesize \ident{scheduled}}=\ident{count(event)}
\end{equation*}
to 
\noindent 
\begin{equation}
N_{\footnotesize \ident{stateTransitions}}=\frac{\ident{max(date(event))}-\ident{min(date(event))}}{\ident{precision}} 
\end{equation}

@@

For numeric types Marlowe supports Integers, while ACTUS is expressed in terms
of real numbers. In order to model a real number in such a setup
we rely on fixed-point precision. The algebra looks like this: 
\begin{verbatim}
  (+)   = AddValue -- x/n + y/n = (x + y)/n
  (-)   = SubValue -- x/n - y/n = (x - y)/n
  a * b = Scale (1 % marloweFixedPoint) $ MulValue a b -- x/n * y/n = (x * y)/n^2
\end{verbatim}
We scale all numbers with \emph{marloweFixedPoint} factor - which only requires modification of \emph{MulValue}.
% SJT this needs more explanation here ^^^
% dk14 I've added some
\noindent
We plan to move Marlowe to fixed-precision numbers for the on-blockchain implementation available later in 2021.

@@

The Marlowe DSL does not support any notion of records,
the variables can only be of type ``Integer''. In order to
map contract state -- \emph{ContractStatePoly} -- we pack a set of Marlowe
variables of type \texttt{Value} and \texttt{Observation}, representing the previous ($t-1$) state, passing them into
\emph{ContractStatePoly}, to apply the polymorphic state transition, and
finally unpack \emph{ContractStatePoly} into a set of Marlowe variables representing the 
 state at $t+1$:
% SJT again I am confused about t-1 to t+1 transition. Why not just one step?  
% dk14. I'll change that. I imagined t as time of event while t+1 is after event 
\noindent 
\begin{equation}
st_{t}=unpack(stateTransition(pack(st_{t-1}))),
\end{equation}
\noindent
where \emph{pack} is a chain of \texttt{UseValue} constructs and \ident{unpack} is a chain of
\texttt{Let}s.
